{\bf Practical Concurrency Testing}

### 2018-10-17 thesis/defense/snakefight.tex
snakefight updates after practice 3 upto bonus slides

	Testing approaches
	\begin{itemize}
		\item Unit tests: only 1 ``most likely'' interleaving
		\item Stress tests: exposes ``less likely'' interleavings (if lucky), % to expose "less likely" ones
			cannot verify
		\item Verification: steep learning curve, unpredictable runtime
			%still no widespread adoption % -- why?
	\end{itemize}
\end{frame}
\begin{frame}{Practical Concurrency Testing}
	%Goals
	\textbf{Landslide}, this thesis's solution
	\begin{itemize}
		\item No randomness, despite nondeterminism
		\item Confirm when you fixed bugs, not just find them % not 'verify'
			% TODO: make alnguage match, "Automatic so and so"
		\item User-friendly enough for students
		% \item Test wide variety of types of concurrency % txn
			% TODO : figure out how to ref 
		% ????
		\item Combines with human intuition to verify tests neither could do alone
	\end{itemize}
	%
	% enter landslide
	% - test yr nondeterministic program without randomness/unpredictability
	% - provide formal verification guarantees so you know when you actually fixed that bug
	%   - (well, subject to assumptions about the test case you're running under, etc etc)
	% - user friendly enough even for ze studence
	% 
	% say: "What if there was a program that... ...I'm here to tell you that I've built such a program,
	% it's called Landslide, and why you should let me graduate for it" (figure out smth more diplomatic here...)
\end{frame}

@@

	Possible interleavings (under given test case) represented as a tree
	\begin{itemize}
		\item Node: Intermediate execution state
		\item Edge: State transition from executing a thread
	\end{itemize}
	\linegap

	State space size depends on which {\bf preemption points} are used.
\end{frame}

\begin{frame}{Challenges} %\& Existing Coping Techniques}
	Unpredictable state space explosion
	\begin{itemize}
		%\item Equivalence reductions: DPOR {\em [Flanagan '05]}, MCR {\em [Huang '15]}
		\item Dynamic Partial Order Reduction (DPOR) {\em [Flanagan '05]}
			\begin{itemize}
				\item identifies \& skips equivalent interleavings
			\end{itemize}
		%\item Heuristic search ordering: ICB {\em [Musuvathi '08]}, PCT {\em [Burckhardt '10]}
		\item Iterative Context Bounding (ICB) {\em [Musuvathi '08]}
			\begin{itemize}
				\item finds bugs faster by prioritizing interleavings with fewer preemptions
			\end{itemize}
			\pause
		\item There will always be a {\bf bigger test case}...
	\end{itemize}
	\pause
	\linegap

	Too much manual intervention %for non-expert use
	\begin{itemize}
		\item State space size estimation {\em [Simsa '12]}
			% TODO: i just switched the order of these; see how that feels
		\item Manual kernel annotations {\em [Blum '12]} %{\tiny that's me!}
			\pause
		\item Widespread adoption requires {\bf fully automated instrumentation}...
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Contributions}
	\hilight{sect-quicksand}{The Quicksand algorithm %combines theory and heuristics to
	practically debugs and verifies many exponential state spaces.}
	\vspace{4.54em}

	\hilight{sect-410}{Landslide's automatic instrumentation makes MC accessible to students.}
	\vspace{4.54em}

	\hilight{sect-htm}{New reduction techniques for transactional memory keep MC up-to-date
	with the latest real-world concurrency techniques.}
	\vspace{4.54em}
\end{frame}
\begin{frame}{Outline}
	\hilight{sect-pastel-quicksand}{The Quicksand algorithm %combines theory and heuristics to
	practically debugs and verifies many exponential state spaces.} \\
	%\hilight{gray}{The Quicksandalgorithm makes MC appropriate for both educational and real-world programs.} \\
	\begin{itemize}
		\item {\bf Quicksand}
		\begin{itemize}
			\item Data-race preemption points \& algorithm overview
			\item Verification \& experimental results
		\end{itemize}
	\end{itemize}
	\hilight{sect-pastel-410}{Landslide's automatic instrumentation makes MC accessible to students.} \\
	%\hilight{gray}{Ad-hoc synchronization heuristics make MC accessible to student programmers.} \\
	\begin{itemize}
		\item {\bf Educational use}
		\begin{itemize}
			\item Pebbles \& Pintos OS projects
			\item Bug-finding, grades, \& surveys
		\end{itemize}
	\end{itemize}

@@

\begin{frame}{Data Race Analysis}
	A {\bf data race} is a pair of memory accesses between two threads, where:
	\begin{itemize}
		\item At least one of the accesses is a write
		\item The threads are not holding the same mutex
		\item Neither access {\em happens-before} the other
			%(e.g., no \texttt{cond\_signal()} in between)
			(e.g., no \texttt{cond\_wait()})
	\end{itemize}
	\pause
	\linegap

	%Prior model checkers hard-code a fixed set of PPs, committing to one state space in advance.
	%Prior model checkers commit to a fixed set of PPs in advance.
	Prior model checkers commit to fixed preemption points in advance.
	\begin{itemize}
		\item ``Preempt on {\tt mutex\_lock()}, {\tt mutex\_unlock()} only''
		\item ``Preempt on every single memory access''
	\end{itemize}
	\linegap

	Insight: Decide where to preempt on-the-fly
	% say: "...then attempt to cope with the resulting exponentially sized problem, however big it may be"

@@

\begin{frame}{Quicksand (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{../../oopsla/dr-jobs-1.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{../../oopsla/dr-jobs-2.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{../../oopsla/dr-jobs-3.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{../../oopsla/dr-jobs-4.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{dr-jobs-maximal.pdf}
	\end{center}
\end{frame}

@@

\begin{frame}{Total Verification}
	\textbf{Convergence theorem:} Testing maximal state space, after detecting all data races, $\equiv$ preempting everywhere!
	\begin{itemize}
		\item $\forall$ arbitrary buggy interleaving, $\exists$ equivalent interleaving % say 'under DPOR' i guess?
			using only sync/data-race PPs, \& Quicksand will eventually reach it.
			\begin{itemize}
				\item ({\em equivalence} defined under DPOR {\em [Flanagan '05]})
			\end{itemize}
	\end{itemize}
	\linegap

@@

\begin{frame}{Setting}
	%CMU (15-410, S'15-S'18), PSU (CMPSC 473, S'18)
	CMU (15-410), PSU (CMPSC 473)
	\begin{itemize}
		\item Thread library (``P2'')
			\begin{itemize}
				\item POSIX-like userspace thread library %(also teams of 2)
				\item Reference ``Pebbles'' kernel provided
					% say: later they implement this in p3
				\item Fixed APIs $\rightarrow$ instrumentation can be automated
				\item Still needs heuristics to handle ad-hoc synchronization
			\end{itemize}
			% FIXME: need introduce p3 here?
			% Or cut?
		\item Kernel (``P3'') - reimplement said reference kernel (CMU only)
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Instrumentation}
	Understanding concurrency primitives/APIs
	\begin{itemize}
		\item Kernel-level testing requires manual annotation {\em [Blum '12]}
		\item P2 (fixed APIs), Pintos (basecode provided) annotated automatically
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Method}
	P2 (CMU + PSU)
	\begin{itemize}
		\item Guest lecture midway during project
		\item Landslide (+ Quicksand) provided to students with 6 test cases %and user guide
		\item Recorded code snapshots and bugs found each use (CMU only)
		\item Surveyed students on usability, false positives, etc.
	\end{itemize}
	\pause
	\linegap

@@

	Bug-finding
	\begin{itemize}
		\item CMU: 115 deterministic, 122 concurrency bugs in 75/103 P2s
		\item U. Chicago: 3 deterministic, 4 concurrency bugs in 3/4 kernels
			% TODO: what to do abt this bullet pts
			%\pause
		\item PSU: (no log collection) %due to IRB issues)
		% \item Case study tej
	\end{itemize}
\end{frame}

\begin{frame}{Effect on Grades (CMU)}
	Compared Landslide users' vs. non-users' submitted P2 grades...
	\begin{itemize}
		\item ``Does Landslide help students submit better projects?''
	\end{itemize}
	\linegap

@@

\begin{frame}{Effect on Grades (CMU)}
	\begin{center}
		\includegraphics[width=\textwidth]{../photo-of-ze-studence.pdf}
	\end{center}
\end{frame}

% TODO: put p values
\begin{frame}{Effect on Grades (CMU)} % - statistically significant?}
	\begin{center}
		\includegraphics[width=\textwidth]{photo-of-ze-studence-cdf-no-pre-study.pdf}
	\end{center}
\end{frame}

\begin{frame}{Effect on Grades (CMU)} % - statistically significant?}
	\begin{center}
		\includegraphics[width=\textwidth]{../photo-of-ze-studence-cdf.pdf}
	\end{center}
\end{frame}

%\begin{frame}{Statistical Significance}
%	Compared %users' vs. non-users'
%	grade distributions with k-sample Anderson-Darling {\em [A., D. '52]}
%	\linegap
%
%	Control 1: non-users from {\bf same semester}
%	\begin{itemize}
%		\item {\bf P2} grade distributions differ with $p = 0.036 < 0.05$
%		\item {\bf P3} grade distributions differ with $p = 0.034 < 0.05$
%			\pause
%		\item Possible bias: stronger students volunteer for Landslide in first place
%	\end{itemize}
%	\pause
%	\linegap
%
%	Control 2: non-users from {\bf past semesters} (when Landslide not offered)
%	\begin{itemize}
%		\item {\bf P2} grade distributions differ with $p = 0.671$...
%		\item {\bf P3} grade distributions differ with $p = 0.145$...
%			\pause
%		\item Possible bias: grading criteria may drift across semesters
%			% FIXME does this bullet belong in this sub list
%		\item Possible bias: maybe Landslide finds bugs the official tests don't
%	\end{itemize}
%\end{frame}

@@

	MCing a transactional program
	\begin{itemize}
		\item Simulate TM, interleaving transactions, rolling-back upon conflicts
			\begin{itemize}
				\item slow; invites state space explosion
			\end{itemize}
		\item Enforce atomicity via scheduler, anticipate aborts on every transaction
			%via analysis
			%\begin{itemize}
			%	\item (under HTM, all transactions might abort)
			%\end{itemize}
		\item Suffices to enumerate all {\em observationally-equivalent} schedules.
	\end{itemize}
\end{frame}

\begin{frame}{Example}
	\begin{center}
		\begin{tabular}{l|l}
			{\bf Thread 1} & {\bf Thread 2} \\
			\hline
			{\tt \hilight{darkorange}{\_xbegin}() == SUCCESS} \\
			{\tt count++;} \\
				& {\tt \hilight{darkorange}{\_xbegin}() == SUCCESS} \\
				& {\tt count++;} \\
				& {\tt \hilight{darkblue}{\_xend}();} \\
			{\tt \hilight{darkblue}{\_xend}()}...? \\
			{\em <conflict detected>} \\
			\\
			{\tt \hilight{darkorange}{\_xbegin}() == ABORTED} \\
			{\tt \hilight{darkorange}{mutex\_lock}(...);} \\
			{\tt count++;} \\
			{\tt \hilight{darkblue}{mutex\_unlock}(...);} \\
		\end{tabular}
	\end{center}
\end{frame}

\begin{frame}{Concurrency model}
	What are a transactional program's {\em observable} behaviours?
	\begin{itemize}
		\item Transactional writes, if successful, appear atomically %to other threads
		\begin{itemize}
			\item Equivalent to a lock protecting that state
				% note that also inserting DR PPs within a transaction is not allowed
		\end{itemize}
			\pause
		\item Simultaneous transactions succeed only if no conflicts
		\begin{itemize}
			\item Equivalent
				%(modulo performance)
				to serialize independent transactions
		\end{itemize}
			\pause
		\item Failing transactions' effects rolled-back with no trace
		\begin{itemize}
			\item Treat each \xbegin as a failure injection point
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

	Equivalence theorem
	\begin{itemize}
		\item Safe to automatically skip all interleavings within transactions %$\qed$
			\begin{itemize}
				\item Exponential state space reduction
				\item Traditional reduction (DPOR) applied on top
			\end{itemize}
		%\item More in bonus slides...
	\end{itemize}
\end{frame}

@@

\breakslide{
\begin{center}
	% TODO: replace with HIL for honban
	\Large Q\&A
	%\includegraphics[width=0.4\textwidth]{../how-i-learned.png}
\end{center}
%\linegap
%
%\begin{center}
%	\includegraphics[width=0.65\textwidth]{3word-questions.png}
%\end{center}
}

@@

\subsection{Quicksand}
\breakslide{\Large \bf \hilight{sect-quicksand}{Bonus Slides - Quicksand}}

@@

\begin{frame}{Quicksand (Iterative Deepening) Reductions}
	If one job is a subset of another, testing either might let us skip the other.
	\begin{center}
		\includegraphics[width=\textwidth]{../../oopsla/reduction-nothing.png}
	\end{center}
\end{frame}
\begin{frame}{Quicksand (Iterative Deepening) Reductions}
	If one job is a subset of another, testing either might let us skip the other.
	\begin{center}
		\includegraphics[width=\textwidth]{../../oopsla/reduction-deferred.png}
	\end{center}
\end{frame}
\begin{frame}{Quicksand (Iterative Deepening) Reductions}
	If one job is a subset of another, testing either might let us skip the other.
	\begin{center}
		\includegraphics[width=\textwidth]{../../oopsla/reduction-complete.png}
	\end{center}
\end{frame}

\begin{frame}{Quicksand results - Bugs found by CPU time}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../quicksand-bugs-hb-talkver.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand results - Bugs found by wall-clock time}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../quicksand-bugs-hb-wallclock-talkver.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand results - Tests fully verified (CPU time)}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../quicksand-verifs-hb-talkver.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand results - MC + DR vs single-pass DR}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../../oopsla/dr-false-negatives-poster.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand results - Representative of real-world programs?}
	Bugs by required number of preemptions to expose:
	\begin{center}
		\footnotesize
	\begin{tabular}{r|c|c}
		{\bf Bound} & {\bf ICB (sync PPs only)} & {\bf ICB (preempt everywhere)} \\
		\hline
		0       & 2     & 1     \\
		1       & 82    & 86    \\
		2       & 16    & 32    \\
		3       & 2     & 3     \\
		4+      & 0     & 0     \\
		\hline
		Total   & 102   & 122   \\
	\end{tabular}
	\end{center}
	\linegap

@@

\begin{frame}{15-410...}
	% TODO:
	% study of tej/etc specific bugs
	% list tests cases provided to studence & PSU-specific ones too
\end{frame}

>>>>

	Testing approaches
	\begin{itemize}
		\item Unit tests: only 1 ``most likely'' interleaving
		\item Stress tests: exposes ``less likely'' interleavings (if lucky), % to expose "less likely" ones
			cannot verify
		\item Verification: steep learning curve, prohibitive runtime
			%still no widespread adoption % -- why?
	\end{itemize}
\end{frame}
\begin{frame}{Practical Concurrency Testing}
	%Goals
	\textbf{Landslide}, this thesis's solution
	\begin{itemize}
		\item No randomness, despite nondeterminism
		\item Confirm when you fixed bugs, not just find them % not 'verify'
			% TODO: make alnguage match, "Automatic so and so"
		\item User-friendly enough for students
		% \item Test wide variety of types of concurrency % txn
			% TODO : figure out how to ref 
		% ????
		\item Combines with human intuition to verify tests neither could do alone
	\end{itemize}
	%
	% enter landslide
	% - test yr nondeterministic program without randomness/unpredictability
	% - provide formal verification guarantees so you know when you actually fixed that bug
	%   - (well, subject to assumptions about the test case you're running under, etc etc)
	% - user friendly enough even for ze studence
	% 
	% say: "What if there was a program that... ...I'm here to tell you that I've built such a program,
	% it's called Landslide, and why you should let me graduate for it" (figure out smth more diplomatic here...)
\end{frame}

@@

	Possible interleavings (under given test case) represented as a tree
	\begin{itemize}
		\item Node: Intermediate execution state
		\item Edge: State transition from executing a thread
		\item Nondeterminism captured as different branches
		\item State space size depends on which {\bf preemption points} used
	\end{itemize}
\end{frame}

\begin{frame}{Challenges} %\& Existing Coping Techniques}
	State space explosion
	\begin{itemize}
		%\item Equivalence reductions: DPOR {\em [Flanagan '05]}, MCR {\em [Huang '15]}
		%\item Dynamic Partial Order Reduction (DPOR) {\em [Flanagan '05]}
		\item Reduction algorithms to skip equivalent interleavings {\em [Flanagan '05]}
			%\begin{itemize}
			%	\item identifies \& skips equivalent interleavings
			%\end{itemize}
		\item Heuristic search ordering to expose bugs faster {\em [Musuvathi '08]}
		%\item Heuristic search ordering: ICB {\em [Musuvathi '08]}, PCT {\em [Burckhardt '10]}
		%\item Iterative Context Bounding (ICB) {\em [Musuvathi '08]}
		%	\begin{itemize}
		%		\item finds bugs faster by prioritizing interleavings with fewer preemptions
		%	\end{itemize}
			\pause
		\item There will always be a {\bf bigger test case}...
	\end{itemize}
	\pause
	\linegap

	Too much manual intervention %for non-expert use
	\begin{itemize}
			% TODO: remember to say, "that lets you choose more intelligently, but still by hand, yr PPs
		\item State space size estimation {\em [Simsa '12]}
			% TODO: i just switched the order of these; see how that feels
		\item Manual kernel annotations {\em [Blum '12]} %{\tiny that's me!}
			\pause
		\item Widespread adoption requires {\bf fully automated instrumentation}...
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Contributions}
	% this vspace used to be 4.54 for all 3 before i rephrased the quicksand bullet
	% even tho it was 2 lines before and 2 lines now
	% i had to do... this??? latex makes no friggin sense
	\hilight{sect-quicksand}{The Quicksand algorithm %combines theory and heuristics to
	practically debugs and verifies many tests despite exponential explosion.}
	\vspace{4.00em}

	\hilight{sect-410}{Landslide's automatic instrumentation makes MC accessible to students.}
	\vspace{4.58em}

	\hilight{sect-htm}{New reduction techniques for transactional memory keep MC up-to-date
	with the latest real-world concurrency techniques.}
	\vspace{4.54em}

\end{frame}
\begin{frame}{Outline}
	\hilight{sect-pastel-quicksand}{The Quicksand algorithm %combines theory and heuristics to
	practically debugs and verifies many tests despite exponential explosion.}
	%\hilight{gray}{The Quicksandalgorithm makes MC appropriate for both educational and real-world programs.} \\
	\begin{itemize}
		\item {\bf Quicksand}
		\begin{itemize}
			\item Data-race preemption points \& algorithm overview
			\item Verification \& experimental results
		\end{itemize}
	\end{itemize}
	\hilight{sect-pastel-410}{Landslide's automatic instrumentation makes MC accessible to students.} \\
	%\hilight{gray}{Ad-hoc synchronization heuristics make MC accessible to student programmers.} \\
	\begin{itemize}
		\item {\bf Educational use}
		\begin{itemize}
			\item Pebbles \& Pintos OS projects
			\item Bug-finding, grades, \& surveys
		\end{itemize}
	\end{itemize}

@@

\begin{frame}{Data Race Analysis {\em [Savage '97]}}
	A {\bf data race} is a pair of memory accesses between two threads, where:
	\begin{itemize}
		\item At least one of the accesses is a write
		\item The threads are not holding the same mutex
		\item Neither access {\em happens-before} the other
		%\item Neither access happens-before the other
			%(e.g., no \texttt{cond\_signal()} in between)
			(e.g., no \texttt{cond\_wait()})
	\end{itemize}
\end{frame}
\begin{frame}{Quicksand}
	%Prior model checkers hard-code a fixed set of PPs, committing to one state space in advance.
	%Prior model checkers commit to a fixed set of PPs in advance.
	Prior model checkers commit to fixed preemption points in advance.
	\begin{itemize}
		\item ``Preempt on {\tt mutex\_lock()}, {\tt mutex\_unlock()} only''
		\item ``Preempt on every single memory access''
	\end{itemize}
	\linegap

	Insight: decide where to preempt on-the-fly
	\begin{itemize}
		\item Data-race analysis identifies new preemption points
		\item User supplies CPU budget
		\item Prioritize based on state space estimation {\em [Simsa '12]}
	\end{itemize}
	% say: "...then attempt to cope with the resulting exponentially sized problem, however big it may be"

@@

\begin{frame}{Quicksand (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{../../oopsla/dr-jobs-1.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{../../oopsla/dr-jobs-2.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	%\includegraphics[width=0.86\textwidth]{../../oopsla/dr-jobs-3.pdf}
	\includegraphics[width=0.86\textwidth]{dr-jobs-3-alt.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{../../oopsla/dr-jobs-4.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand (in Pictures)}
	\begin{center}
	\vspace{-0.88em}
	\includegraphics[width=0.86\textwidth]{dr-jobs-maximal.pdf}
	\end{center}
\end{frame}

@@

\begin{frame}{Total Verification}
	\textbf{Convergence theorem:} Testing maximal state space, after detecting all data races, $\equiv$ preempting everywhere!
	\begin{itemize}
		\item $\forall$ arbitrary buggy interleaving, $\exists$ equivalent interleaving % say 'under DPOR' i guess?
			using only sync/data-race PPs, \& Quicksand will eventually reach it.
			%\begin{itemize}
			%	\item ({\em equivalence} defined under DPOR {\em [Flanagan '05]})
			%\end{itemize}
	\end{itemize}
	\linegap

@@

\begin{frame}{Setting}
	%CMU (15-410, S'15-S'18), PSU (CMPSC 473, S'18)
	CMU (15-410), PSU (CMPSC 473)
	\begin{itemize}
		\item Thread library (``P2'')
			\begin{itemize}
				\item POSIX-like userspace thread library %(also teams of 2)
				\item Implement {\tt thr\_create()}/{\tt exit()}/{\tt join()}, {\tt mutex\_lock()}, etc.
				\item Reference ``Pebbles'' kernel provided
					% say: later they implement this in p3
				%\item Fixed APIs $\rightarrow$ instrumentation can be automated
				%\item Still needs heuristics to handle ad-hoc synchronization
			\end{itemize}
		\item Kernel (``P3'') - reimplement said reference kernel (CMU only)
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Instrumentation}
	Understanding concurrency primitives/APIs
	\begin{itemize}
		\item Testing Pebbles kernels requires manual annotation {\em [Blum '12]}
		\item P2 (fixed APIs), Pintos (basecode provided) annotated automatically
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{User Study}
	P2 (CMU + PSU)
	\begin{itemize}
		\item Guest lecture midway during project
		\item Landslide (+ Quicksand) provided to students with 6 test cases %and user guide
		\item Recorded code snapshots and bugs found each use (CMU only)
		\item Surveyed students on usability, false positives, etc.
	\end{itemize}
	\pause
	\linegap

@@

	Bug-finding
	\begin{itemize}
		\item CMU: 115 deterministic, 122 concurrency bugs in 75/103 P2s
			%\begin{itemize}
			%	\item 3 false positives (self-reported)
			%\end{itemize}
		\item U. Chicago: 3 deterministic, 4 concurrency bugs in 3/4 kernels
			% TODO: what to do abt this bullet pts
			%\pause
		\item PSU: (no log collection) %due to IRB issues)
		% \item Case study tej
	\end{itemize}
\end{frame}

\begin{frame}{Effect on Grades (CMU)}
	Compared Landslide users' vs. non-users' submitted P2 grades
	\begin{itemize}
		\item ``Does Landslide help students submit better projects?''
	\end{itemize}
	\linegap

@@

\begin{frame}{Effect on Grades (CMU)}
	\begin{center}
		\includegraphics[width=\textwidth]{../photo-of-ze-studence.pdf}
	\end{center}
\end{frame}
\begin{frame}{Effect on Grades (CMU)} % - statistically significant?}
	\begin{center}
		\includegraphics[width=\textwidth]{photo-of-ze-studence-cdf-no-pre-study.pdf}
	\end{center}
\end{frame}
\begin{frame}{Effect on Grades (CMU)} % - statistically significant?}
	\begin{center}
		\includegraphics[width=\textwidth]{photo-of-ze-studence-cdf.pdf}
	\end{center}
\end{frame}

@@

	MCing a transactional program
	\begin{itemize}
		\item Simulate TM, interleaving transactions, rolling-back upon conflicts
			\begin{itemize}
				\item Slow; invites state space explosion
			\end{itemize}
		\item Enforce atomicity via scheduler, anticipate aborts on every transaction
			\vspace{-1em} % latex... WTF??
			\begin{itemize}
				\item Suffices to enumerate all observationally-equivalent schedules.
			\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Example (Hardware POV)}
	\begin{center}
		\begin{tabular}{l|l}
			{\bf Thread 1} & {\bf Thread 2} \\
			\hline
			{\tt \hilight{darkorange}{\_xbegin}() == SUCCESS} \\
			{\tt count++;} \\
				& {\tt \hilight{darkorange}{\_xbegin}() == SUCCESS} \\
				& {\tt count++;} \\
				& {\tt \hilight{darkblue}{\_xend}();} \\
			{\tt \hilight{darkblue}{\_xend}()}...? \\
			{\em <conflict detected>} \\
			\\
			{\tt \hilight{darkorange}{\_xbegin}() == ABORTED} \\
			{\tt \hilight{darkorange}{mutex\_lock}(...);} \\
			{\tt count++;} \\
			{\tt \hilight{darkblue}{mutex\_unlock}(...);} \\
		\end{tabular}
	\end{center}
\end{frame}

\begin{frame}{Concurrency model}
	What are a transactional program's {\em observable} behaviours?
	\begin{itemize}
		\item Transactional writes, if successful, appear atomically %to other threads
		%\begin{itemize}
		%	\item Equivalent to a lock protecting that state
		%		% note that also inserting DR PPs within a transaction is not allowed
		%\end{itemize}
		%	\pause
		\item Simultaneous transactions succeed only if no conflicts
		%\begin{itemize}
		%	\item Equivalent
		%		%(modulo performance)
		%		to serialize independent transactions
		%\end{itemize}
		%	\pause
		\item Failing transactions' effects rolled-back with no trace
		%\begin{itemize}
		%	\item Treat each \xbegin as a failure injection point
		%\end{itemize}
	\end{itemize}
	\pause
	\linegap

	Equivalence theorem
	\begin{itemize}
		\item $\forall$ interleaving within transactions, $\exists$ equivalent interleaving
			treating them atomically, which either succeeds or aborts immediately
		\item Safe to automatically skip all interleavings within transactions %$\qed$
			\begin{itemize}
				\item Exponential state space reduction
				\item Traditional reduction %(DPOR)
					applied on top
			\end{itemize}
		%\item More in bonus slides...
	\end{itemize}
\end{frame}

@@

\breakslide{
\begin{center}
	\includegraphics[width=0.4\textwidth]{../how-i-learned.png}
\end{center}
}

@@

\subsection{Prior Work}
\breakslide{\Large \bf Bonus Slides - Prior Work}

\begin{frame}{Dynamic Partial Order Reduction (DPOR) {\em [Flanagan '05]}}
	State space for 2 threads doing {\tt tmp = count; tmp++; count = tmp;} (fixing one thread to run first)
	\begin{center}
		\includegraphics[width=\textwidth]{../statespace-tree.pdf}
	\end{center}
	\linegap

	Number of ways to interleave $M$ and $N$ events: ${M+N \choose M}$
\end{frame}
%\begin{frame}{DPOR example}
\begin{frame}{Dynamic Partial Order Reduction (DPOR) {\em [Flanagan '05]}}
	Result of a single iteration of DPOR: alternate branches
	which would reorder memory conflicts are ``tagged'' for exploration
	\begin{center}
		\includegraphics[width=\textwidth]{../dpor-example-0.pdf}
	\end{center}
\end{frame}
\begin{frame}{Dynamic Partial Order Reduction (DPOR) {\em [Flanagan '05]}}
	Result after 3 iterations; D-J pruned
	\begin{center}
		\includegraphics[width=\textwidth]{../dpor-example-1.pdf}
	\end{center}
\end{frame}
\begin{frame}{Dynamic Partial Order Reduction (DPOR) {\em [Flanagan '05]}}
	Right subtree symmetric; pruning L+M possible with {\em sleep set reduction}
	\begin{center}
		\includegraphics[width=0.9\textwidth]{../dpor-example-2.pdf}
	\end{center}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quicksand}
\breakslide{\Large \bf \hilight{sect-quicksand}{Bonus Slides - Quicksand}}

\begin{frame}{Quicksand Reductions}
	If one job is a subset of another, testing either might let us skip the other.
	\begin{center}
		\includegraphics[width=\textwidth]{../../oopsla/reduction-nothing.png}
	\end{center}
\end{frame}
\begin{frame}{Quicksand Reductions}
	If one job is a subset of another, testing either might let us skip the other.
	\begin{center}
		\includegraphics[width=\textwidth]{../../oopsla/reduction-deferred.png}
	\end{center}
\end{frame}
\begin{frame}{Quicksand Reductions}
	If one job is a subset of another, testing either might let us skip the other.
	\begin{center}
		\includegraphics[width=\textwidth]{../../oopsla/reduction-complete.png}
	\end{center}
\end{frame}

\begin{frame}{Pure HB vs Limited HB - Bugs found by CPU time}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../quicksand-bugs-hb-talkver.pdf}
	\end{center}
\end{frame}
\begin{frame}{Pure HB vs Limited HB - Bugs found by wall-clock time}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../quicksand-bugs-hb-wallclock-talkver.pdf}
	\end{center}
\end{frame}
\begin{frame}{Pure HB vs Limited HB - Tests fully verified (CPU time)}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../quicksand-verifs-hb-talkver.pdf}
	\end{center}
\end{frame}
\begin{frame}{Pure HB vs Limited HB - MC + DR vs single-pass DR}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{../../oopsla/dr-false-negatives-poster.pdf}
	\end{center}
\end{frame}
\begin{frame}{Quicksand results - Representative of real-world programs?}
	Bugs by required number of preemptions to expose:
	\begin{center}
		\footnotesize
	\begin{tabular}{r|c|c}
		{\bf Bound} & {\bf ICB (sync PPs only)} & {\bf ICB (preempt everywhere)} \\
		\hline
		0       & 2     & 1     \\
		1       & 82    & 86    \\
		2       & 16    & 32    \\
		3       & 2     & 3     \\
		4+      & 0     & 0     \\
		\hline
		Total   & 102   & 122   \\
	\end{tabular}
	\end{center}
	\linegap

@@

\begin{frame}{Statistical Significance}
	Compared %users' vs. non-users'
	grade distributions with k-sample Anderson-Darling {\em [A., D. '52]}
	\linegap

	Control 1: non-users from {\bf same semester}
	\begin{itemize}
		\item {\bf P2} grade distributions differ with $p = 0.036 < 0.05$
		\item {\bf P3} grade distributions differ with $p = 0.034 < 0.05$
			\pause
		\item Possible bias: stronger students volunteer for Landslide in first place
	\end{itemize}
	\pause
	\linegap

	Control 2: non-users from {\bf past semesters} (when Landslide not offered)
	\begin{itemize}
		\item {\bf P2} grade distributions differ with $p = 0.671$...
		\item {\bf P3} grade distributions differ with $p = 0.145$...
			\pause
		\item Possible bias: grading criteria may drift across semesters
			% FIXME does this bullet belong in this sub list
		\item Possible bias: maybe Landslide finds bugs the official tests don't
	\end{itemize}
\end{frame}
